---
title: "Covid91 vaccine study Final2021"
author: "Nicole Orizino"
date: "April 2021"
output: 
  html_document:
    number_sections: true
    toc: true

---

```{r include=FALSE,echo=FALSE}
require(tidyverse)
require(tigerstats)
require(forcats)
OldData <-read.csv(file="covid91.csv",header=TRUE)
#recode to remove the annoying "X" from the dataframe
FinalData <- OldData %>% mutate(infected = recode(infected, aids = "covid", none = "OK" )) %>% select(-one_of("X"))
```

# Introduction

A new  preventative vaccine (covid91) is under test in clinical trials. Given the nature of this vaccine it is suspected that the drug may exhibit different effectiveness across the variety of subpopulations tracked. Consider population subgroups consisting of Males, Females, just the drug users, and just those in the LGBTQ community. Write a complete report that separately considers this drug for each subgroup and your recommendations as to whether people in the subgroup should take the vaccine. Structure your report in four sections each devoted to one of the four subgroups, use null and alternative hypothesis statements and correct statistical verbage when referencing p-values in conclusions.

# Methoods 

The first step is, isolating the subpopulations
```{r}
maledata<- filter(FinalData,sex=="guy")
maledata$sex <- as.factor(maledata$sex)
maledata$infected=factor(maledata$infected)
maledata$infected<-relevel(maledata$infected, "OK")
```

```{r}
femaledata<- filter(FinalData, sex=="gal")
femaledata$sex <- as.factor(femaledata$sex)
femaledata$infected=factor(femaledata$infected)
femaledata$infected<-relevel(femaledata$infected, "OK")
levels(femaledata$infected)
```

```{r}
Druggiedata<- filter(FinalData, DrugUser=="yes")
Druggiedata$DrugUser <- as.factor(Druggiedata$DrugUser)
Druggiedata$infected=factor(Druggiedata$infected)
Druggiedata$infected<-relevel(Druggiedata$infected, "OK")
```

```{r}
LGBTQdata<- filter(FinalData, LGBTQ=="gay")
LGBTQdata$LGBTQ <- as.factor(LGBTQdata$LGBTQ)
LGBTQdata$infected=factor(LGBTQdata$infected)
LGBTQdata$infected<-relevel(LGBTQdata$infected, "OK")
```
# Males

## Bar Chart


```{r}
Males <- maledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Males,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Males,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 340 Males who were infected with covid-19, 57% were given the drug, and 43% were given the placebo.

## Numerical Summary
```{r}
table1 <- xtabs(~treatment + infected, data=maledata)
rowPerc(table1)
colPerc(table1)
print(table1)
```

The table shows that about 340 Males were infected with Covid-19 out of approximately 10,400 Males in this study. The table here is showing that 3.72 percent of Males who received the drug got Covid, and 2.55 percent of the Males who got the placebo contracted Covid-19, this suggest that the vaccine is not effective for Males.

## Inferential Results

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

The Chi Squared table shows that the P-Value of 0.0006181 is so small that it is highly unlikely that the results are by chance. 
$H_0$ the null hypothesis given taking the vaccine there is no difference of weather or not a Male contract Covid-19 or not. 
$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in weather or not a Male contract Covid-19. Therefore we Reject the Null Hypothesis. 

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odds ratio and computing it into a percent of effectiveness by usung the formula 0/(1+0)*100 you get 59.5% effectiveness.

Overall the vaccine is 40% effective in the whole population, which means it is absolutely ineffective. 0.678/(1+0.678)*100 = 40.40524 or 40%.

# Females

## Bar Chart

```{r}
Females <- femaledata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Females,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Females,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 340 Males who were infected with covid-19, 57% were given the drug, and 43% were given the placebo.

## Numerical Summary
```{r}
table2 <- xtabs(~infected + treatment, data=femaledata)
rowPerc(table2)
colPerc(table2)
print(table2)
```

The table shows that about 340 Males were infected with Covid-19 out of approximately 10,400 Males in this study. The table here is showing that 3.72 percent of Males who received the drug got Covid, and 2.55 percent of the Males who got the placebo contracted Covid-19, this suggest that the vaccine is not effective for Males.


## Inferential Results

```{r}
chisq.test(table2)
chisqtestGC(table2)
fisher.test(table2)
```

The Chi Squared table shows that the P-Value of 0.0006181 is so small that it is highly unlikely that the results are by chance. 
$H_0$ the null hypothesis given taking the vaccine there is no difference of weather or not a Male contract Covid-19 or not. 
$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in weather or not a Male contract Covid-19. Therefore we Reject the Null Hypothesis. 

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odds ratio and computing it into a percent of effectiveness by usung the formula 0/(1+0)*100 you get 59.5% effectiveness.


# Drug Users

## Bar Chart


```{r}
Druggies <- Druggiedata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(Druggies,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(Druggies,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 340 Males who were infected with covid-19, 57% were given the drug, and 43% were given the placebo.

## Numerical Summary
```{r}
table3 <- xtabs(~treatment + infected, data=Druggiedata)
rowPerc(table3)
colPerc(table3)
print(table3)
```

The table shows that about 340 Males were infected with Covid-19 out of approximately 10,400 Males in this study. The table here is showing that 3.72 percent of Males who received the drug got Covid, and 2.55 percent of the Males who got the placebo contracted Covid-19, this suggest that the vaccine is not effective for Males.

## Inferential Results

```{r}
chisq.test(table3)
chisqtestGC(table3)
fisher.test(table3)
```

The Chi Squared table shows that the P-Value of 0.0006181 is so small that it is highly unlikely that the results are by chance. 
$H_0$ the null hypothesis given taking the vaccine there is no difference of weather or not a Male contract Covid-19 or not. 
$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in weather or not a Male contract Covid-19. Therefore we Reject the Null Hypothesis. 

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odds ratio and computing it into a percent of effectiveness by usung the formula 0/(1+0)*100 you get 59.5% effectiveness.

Overall the vaccine is 40% effective in the whole population, which means it is absolutely ineffective. 0.678/(1+0.678)*100 = 40.40524 or 40%.


# LGBTQ

## Bar Chart


```{r}
LGBTQ <- LGBTQdata %>% group_by(infected,treatment) %>% summarize(count=n()) %>% mutate(prcnt=count/sum(count))


# the group_by followed by summarize(count=n())

basicC <- ggplot(LGBTQ,aes(x=infected,y=count,fill=treatment))
basicC + geom_bar(stat="identity",position="dodge")

#Now for percentage plot

basicCC <- ggplot(LGBTQ,aes(x=infected,y=prcnt*100,fill=treatment)) 
basicCC + geom_bar(stat="identity", position = "dodge")
```

This barchart shows that out of the approximately 340 Males who were infected with covid-19, 57% were given the drug, and 43% were given the placebo.

## Numerical Summary
```{r}
table4 <- xtabs(~treatment + infected, data=LGBTQdata)
rowPerc(table4)
colPerc(table4)
print(table4)
```

The table shows that about 340 Males were infected with Covid-19 out of approximately 10,400 Males in this study. The table here is showing that 3.72 percent of Males who received the drug got Covid, and 2.55 percent of the Males who got the placebo contracted Covid-19, this suggest that the vaccine is not effective for Males.

## Inferential Results

```{r}
chisq.test(table4)
chisqtestGC(table4)
fisher.test(table4)
```

The Chi Squared table shows that the P-Value of 0.0006181 is so small that it is highly unlikely that the results are by chance. 
$H_0$ the null hypothesis given taking the vaccine there is no difference of weather or not a Male contract Covid-19 or not. 
$H_A$ the alternative hypothesis is that the vaccine in fact does make a difference in weather or not a Male contract Covid-19. Therefore we Reject the Null Hypothesis. 

The odds ratio from the Fisher Exact test is 1.47:1, when taking the odds ratio and computing it into a percent of effectiveness by usung the formula 0/(1+0)*100 you get 59.5% effectiveness.

Overall the vaccine is 40% effective in the whole population, which means it is absolutely ineffective. 0.678/(1+0.678)*100 = 40.40524 or 40%.


# Discustions/Conclutions

